\LARGE \bf Mean Estimation from One-Bit Measurements

### 2022-04-20 proof-sgd.tex
addressed specific commets in R2

It remains to show that the average $\bar{\theta}_n$ is regular \newtext{at $\theta$ with the limiting distribution $\Ncal(0,(4f(0)^2)^{-1})$}:
\begin{lemma}
  \label{lemma:sgd-median-regular}
  Let $h_n \to h \in \R$, and define $P_{n,h} = P_{\theta + h_n / \sqrt{n}}^n$.
  Then 
  \begin{align}
    \label{eqn:sgd-median-regular}
    \newtext{
    \sqrt{n}\left( \bar{\theta}_n - \theta \right)
    \mathop{\cd}_{P_{n,h}}
    \normal\left( h,\frac{1}{4 f(0)^2}\right).
    }
  \end{align}
\end{lemma}
\begin{proof}
  To show the convergence~\eqref{eqn:sgd-median-regular} we use the
  following refinement of Corollary~\ref{corollary:polyak-juditsky}, which
  provides a generalized convergence result for iteratively defined
  $\theta_n$, and whose
  proof we defer to Section~\ref{proof:normal-expansion}.
  %
  \begin{corollary}
    \label{corollary:normal-expansion}
    Let the conditions of Corollary~\ref{corollary:polyak-juditsky} hold,
    meaning that $\theta_i = \theta_{i-1} + \gamma_i \varphi(X_i -
    \theta_{i-1})$ for $X_i = \theta + Z_i$, where $\{Z_i\}$ are i.i.d.\ with
    $\E[\varphi(Z)] = 0$. Additionally assume the local smoothness
    condition that there exist $0 < \lambda \le 1$ and $K < \infty$ such
    that
    \begin{equation}
      \label{eqn:additional-local-smooth}
      \E[|\varphi(x + Z_1) - \varphi(Z)|^2]
      \le K (|x|^\lambda + x^2).
    \end{equation}
    Set $\Delta_i \defeq \theta_i - \theta$ and $\bar{\Delta}_n \defeq \frac{1}{n}
    \sum_{i=1}^n \Delta_i$. Then
    \begin{enumerate}[(i)]
    \item \label{item:regularity}
      The sequence $\{\Delta_i\}$ is regular, that is,
      \begin{equation}
        \sqrt{n} \bar{\Delta}_n
        = -\frac{1}{\sqrt{n}} \frac{1}{\psi'(0)} \sum_{i=1}^{n-1} \varphi(Z_i)
        + o_{P,n}(1).
        \label{eq:normal_expansion_lem}
      \end{equation}
    \item \label{item:apply-le-cam} Let $\{Z_i\}$ as in
      Corollary~\ref{corollary:polyak-juditsky} have absolutely continuous
      density $p$ with median $0$, define $\score_h(z) = \frac{p'(z - h)}{p(z
        - h)}$, and assume that $I_h \defeq \E_p[\score_h(Z_1)^2]$ is
      continuous in $h$ near 0.  Then for any converging sequence $h_n \to h$,
      \begin{equation*}
        \sqrt{n} \bar{\Delta}_n
        \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}^n}
        \normal\left( \frac{-h}{\psi'(0)} \E_p[\varphi(Z_1) \score_0(Z_1)],
        \frac{\chi(0)}{\psi'^2(0)} \right).
      \end{equation*}
    \end{enumerate}
  \end{corollary}

  We now verify that the setting of Theorem~\ref{thm:sgd}
  (and Lemma~\ref{lemma:sgd-median-regular}) satisfies the
  conditions of Corollary~\ref{corollary:polyak-juditsky}. First, we have
  the obvious fact that
  \begin{equation*}
    |\sgn(z) - \sgn(x + z)|
    \le 2 \cdot \indic{|x| \ge |z|}.
  \end{equation*}
  Recalling that the density $f$ is Lipschitz with median 0,
  for $\varphi(z) = \sgn(z)$, and $Z = X - \theta$ distributed with
  density $f$, we have
  \begin{align*}
    \ex{ \left| \varphi(Z) - \varphi(x + Z) \right| }
    \leq 2 \Prob\left( |Z_1| \le |x|  \right)
    = 2 \int_{-|x|}^{|x|} f(t) dt
    \le 4 f(0)|x| + 2 \int_{-|x|}^{|x|}
    \lip(f) t dt
    = 4 f(0) |x| + 2 \lip(f) x^2
  \end{align*}
  where $\lip(f)$ is the Lipschitz constant of $f$.
\newtext{It follows that condition~\eqref{eqn:additional-local-smooth} holds.}  
   %For large $x$ we always have  $\E[|\varphi(Z) - \varphi(x + Z)|] \le 2$, so that condition~\eqref{eqn:additional-local-smooth} holds.
  In addition, we have
  \begin{align*}
\newtext{ \E_p[\varphi(Z_1) \score_0(Z_1)] } & = 
    \int_{\R} \varphi(x) f'(x ) dx
    = \int_{\R} \sgn(x) f'(x ) dx\\
     &  =
    \int_0^\infty f'(x ) dx
    - \int_{-\infty}^0 f'(x ) dx
    = -2 f(0) = -\psi'(0).
  \end{align*}
  Corollary~\ref{corollary:normal-expansion}
  now implies the convergence \eqref{eqn:sgd-median-regular}.
\end{proof}

@@

  The term \eqref{eq:PJ_proof2} is somewhat more challenging to control.
  We define
  \begin{equation*}
    \epsilon_i \defeq \psi(\Delta_{i-1}) + \varphi(Z_i)
    - \varphi(\Delta_{i-1}+Z_i),
  \end{equation*}
  and let
  $\mc{F}_i = \sigma(Z_1, \ldots, Z_i)$ be the $\sigma$-field of the randomness
  through time $i$. We use a square integrable martingale convergence
  theorem~\cite[Exercise~5.3.35]{Dembo16}. Noting that
  $\Delta_i \in \mc{F}_i$, we have
  \begin{align}
    \E[\epsilon_i^2 \mid \mc{F}_{i-1}]
    & = \E[(\psi(\Delta_{i-1}) + \varphi(Z_i)
      - \varphi(\Delta_{i-1} + Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \le 2 \psi(\Delta_{i-1})^2
    + 2 \E[(\varphi(\Delta_{i-1} + Z_i) - \varphi(Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \le
    K \left[|\Delta_{i-1}|^{1 + \lambda}
      + |\Delta_{i-1}|^\lambda
      + \Delta_{i-1}^2 \right],
    \label{eqn:bound-psi-error-expectations}
  \end{align}
  where inequality~\eqref{eqn:bound-psi-error-expectations} follows by the
  conditions~\eqref{eqn:local-hessian-psi}
  and~\eqref{eqn:additional-local-smooth}, and $\E[\varepsilon_i \mid
    \mc{F}_{i-1}] = 0$ for all $i$ by definition of $\psi(x) = \E[\varphi(x
    + Z)]$ and that $\psi(0) = 0$.  We now control the expectations of these
  quantities. 
  \newtext{
  As in the proof of Theorems~2 and~4 in \cite{polyak1992acceleration}, the Robbins-Siegmund Theorem \cite{robbins1971convergence} applies to the increment of $|\Delta_t|^2$ implies that for every 
  $\epsilon>0$ there exists some $R>0$ such that
  \begin{align}
  \label{eq:sup_delta}
  \Prob\left(\sup_i |\Delta_i| \le R \right) \ge 1-\epsilon.
  \end{align} 
  Define the the stopping time $\tau_R \defeq \inf
  \{i : |\Delta_i| > R\}$, which satisfies $\{\tau_R \le i\} \in \mc{F}_i$
  for each $i$. 
  }
  Then using~\cite[Eq.~(A14)]{polyak1992acceleration}, we have
  \begin{equation*}
    \E[\Delta_i^2 \indic{\tau_R > i}] \le K \gamma_i,
  \end{equation*}
  and so inequality~\eqref{eqn:bound-psi-error-expectations} gives that
  \begin{equation*}
    \E\bigg[\sum_{i = 1}^\infty \frac{1}{i}
      |\varepsilon_i|^2 \indic{\tau_R > i} \bigg]
    \le K \sum_{i = 1}^\infty \frac{\gamma_i^\lambda}{i}
    < \infty
    ~~ \mbox{so} ~~
    \sum_{i = 1}^\infty \frac{1}{i}
    \varepsilon_i^2 \indic{\tau_R > i} < \infty
    ~ \mbox{a.s.}
  \end{equation*}
  by Condition~\eqref{eqn:lazy-gamma}. Since
  \begin{align*}
\left\{ \sum_{i=1}^\infty \frac{|\Delta_i|^{1+\lambda}}{\sqrt{i}} < \infty \subset \left\{ \sup_i |\Delta_i| \leq R \right\} \cap \left\{ \sum_{i=0}^\infty  \right\}
 \right\},
  \end{align*}
  \newtext{it follows from \eqref{eq:sup_delta} and by the arbitrary choice of $\epsilon$ that}
  %\newtext{As it follows from \cite[Eq.~(A11)]{polyak1992acceleration} that $\Prob\left( \sup |\Delta_t| < \infty  \right)=1$}, we obtain
  %\newtext{As Lemma~\ref{lemma:converging-power-sum} implies in particular that $\{\Delta_i\}$ is a.s. bounded, with probability 1 there exists some (random) $R < \infty$ such that $\tau_R = \infty$.} We obtain that 
  \begin{equation*}
    \sum_{i = 1}^\infty \frac{1}{i}
    \varepsilon_i^2  < \infty
    ~ \mbox{a.s.}.
  \end{equation*}
  Applying the square integrable martingale convergence
  theorem of \cite[Ex.~5.3.35]{Dembo16}, we have
  \begin{equation*}
    \frac{1}{\sqrt{n}} \sum_{i = 1}^n
    \left(\frac{1}{\psi'(0)} + w_i^n\right) \epsilon_i \cas 0,
  \end{equation*}
  so that both equations~\eqref{eq:PJ_proof1} and~\eqref{eq:PJ_proof2}
  converge almost surely to 0.
\end{proof}

>>>>

It remains to show that the average $\bar{\theta}_n$ is regular \newtext{at $\theta$ with the limiting distribution $\Ncal(0,(4f(0)^2)^{-1})$}:
\begin{lemma}
  \label{lemma:sgd-median-regular}
  Let $h_n \to h \in \R$, and define $P_{n,h} = P_{\theta + h_n / \sqrt{n}}^n$.
  Then 
  \begin{align}
    \label{eqn:sgd-median-regular}
    \newtext{
    \sqrt{n}\left( \bar{\theta}_n - \theta \right)
    \mathop{\cd}_{P_{n,h}}
    \normal\left( h,\frac{1}{4 f(0)^2}\right).
    }
  \end{align}
\end{lemma}
\begin{proof}
  To show the convergence~\eqref{eqn:sgd-median-regular} we use the
  following refinement of Corollary~\ref{corollary:polyak-juditsky}, which
  provides a generalized convergence result for iteratively defined
  $\theta_n$, and whose
  proof we defer to Section~\ref{proof:normal-expansion}.
  %
  \begin{corollary}
    \label{corollary:normal-expansion}
    Let the conditions of Corollary~\ref{corollary:polyak-juditsky} hold,
    meaning that $\theta_i = \theta_{i-1} + \gamma_i \varphi(X_i -
    \theta_{i-1})$ for $X_i = \theta + Z_i$, where $\{Z_i\}$ are i.i.d.\ with
   $\E[Z_1]=0$ and  $\E[\varphi(Z_1)] = 0$. Additionally assume the local smoothness
    condition that there exist $0 < \lambda \le 1$ and $K < \infty$ such
    that
    \begin{equation}
      \label{eqn:additional-local-smooth}
      \E[|\varphi(x + Z_1) - \varphi(Z_1)|^2]
      \le K (|x|^\lambda + x^2).
    \end{equation}
    Set $\Delta_i \defeq \theta_i - \theta$ and $\bar{\Delta}_n \defeq \frac{1}{n}
    \sum_{i=1}^n \Delta_i$. Then
    \begin{enumerate}[(i)]
    \item \label{item:regularity}
      The sequence $\{\Delta_i\}$ is regular, that is,
      \begin{equation}
        \sqrt{n} \bar{\Delta}_n
        = -\frac{1}{\sqrt{n}} \frac{1}{\psi'(0)} \sum_{i=1}^{n-1} \varphi(Z_i)
        + o_{P,n}(1).
        \label{eq:normal_expansion_lem}
      \end{equation}
    \item \label{item:apply-le-cam} Let $\{Z_i\}$ as in
      Corollary~\ref{corollary:polyak-juditsky} have absolutely continuous
      density $p$ with median $0$, define $\score_h(z) = \frac{p'(z - h)}{p(z
        - h)}$, and assume that $I_h \defeq \E_p[\score_h(Z_1)^2]$ is
      continuous in $h$ near 0.  Then for any converging sequence $h_n \to h$,
      \begin{equation*}
        \sqrt{n} \bar{\Delta}_n
        \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}^n}
        \normal\left( \frac{-h}{\psi'(0)} \E_p[\varphi(Z_1) \score_0(Z_1)],
        \frac{\chi(0)}{\psi'^2(0)} \right).
      \end{equation*}
    \end{enumerate}
  \end{corollary}

  We now verify that the setting of Theorem~\ref{thm:sgd}
  (and Lemma~\ref{lemma:sgd-median-regular}) satisfies the
  conditions of Corollary~\ref{corollary:normal-expansion}. First, we have
  the obvious fact that
  \begin{equation*}
    |\sgn(z) - \sgn(x + z)|
    \le 2 \cdot \indic{|x| \ge |z|}.
  \end{equation*}
  Recalling that the density $f$ is Lipschitz with median 0,
  for $\varphi(z) = \sgn(z)$, and $Z = X - \theta$ distributed with
  density $f$, we have
  \begin{align*}
    \ex{ \left| \varphi(Z) - \varphi(x + Z) \right| }
    \leq 2 \Prob\left( |Z_1| \le |x|  \right)
    = 2 \int_{-|x|}^{|x|} f(t) dt
    \le 4 f(0)|x| + 2 \int_{-|x|}^{|x|}
    \lip(f) t dt
    = 4 f(0) |x| + 2 \lip(f) x^2
  \end{align*}
  where $\lip(f)$ is the Lipschitz constant of $f$.
\newtext{It follows that condition~\eqref{eqn:additional-local-smooth} holds.}  
   %For large $x$ we always have  $\E[|\varphi(Z) - \varphi(x + Z)|] \le 2$, so that condition~\eqref{eqn:additional-local-smooth} holds.
  In addition, we have
  \begin{align*}
\newtext{ \E_p[\varphi(Z_1) \score_0(Z_1)] } & = 
    \int_{\R} \varphi(x) f'(x ) dx
    = \int_{\R} \sgn(x) f'(x ) dx\\
     &  =
    \int_0^\infty f'(x ) dx
    - \int_{-\infty}^0 f'(x ) dx
    = -2 f(0) = -\psi'(0).
  \end{align*}
  Corollary~\ref{corollary:normal-expansion}
  now implies the convergence \eqref{eqn:sgd-median-regular}.
\end{proof}

@@

  The term \eqref{eq:PJ_proof2} is somewhat more challenging to control.
  We define
  \begin{equation*}
    \epsilon_i \defeq \psi(\Delta_{i-1}) + \varphi(Z_i)
    - \varphi(\Delta_{i-1}+Z_i),
  \end{equation*}
  and let
  $\mc{F}_i = \sigma(Z_1, \ldots, Z_i)$ be the $\sigma$-field of the randomness
  through time $i$. We use a square integrable martingale convergence
  theorem~\cite[Exercise~5.3.35]{Dembo16}. Noting that
  $\Delta_i \in \mc{F}_i$, we have
  \begin{align}
    \E[\epsilon_i^2 \mid \mc{F}_{i-1}]
    & = \E[(\psi(\Delta_{i-1}) + \varphi(Z_i)
      - \varphi(\Delta_{i-1} + Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \le 2 \psi(\Delta_{i-1})^2
    + 2 \E[(\varphi(\Delta_{i-1} + Z_i) - \varphi(Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \le
    K \left[|\Delta_{i-1}|^{1 + \lambda}
      + |\Delta_{i-1}|^\lambda
      + \Delta_{i-1}^2 \right],
    \label{eqn:bound-psi-error-expectations}
  \end{align}
  where inequality~\eqref{eqn:bound-psi-error-expectations} follows by the
  conditions~\eqref{eqn:local-hessian-psi}
  and~\eqref{eqn:additional-local-smooth}, and $\E[\varepsilon_i \mid
    \mc{F}_{i-1}] = 0$ for all $i$ by definition of $\psi(x) = \E[\varphi(x
    + Z)]$ and that $\psi(0) = 0$.  We now control the expectations of these
  quantities. For $R<\infty$, define the the stopping time $\tau_R \defeq \inf
  \{i : |\Delta_i| > R\}$, which satisfies $\{\tau_R \le i\} \in \mc{F}_i$
  for each $i$. Then using~\cite[Eq.~(A13-A14)]{polyak1992acceleration}, we have
  \begin{equation*}
    \E[\Delta_i^2 \indic{\tau_R > i}] \le K \gamma_i,
  \end{equation*}
  and so inequality~\eqref{eqn:bound-psi-error-expectations} gives that
  \begin{equation*}
    \E\bigg[\sum_{i = 1}^\infty \frac{1}{i}
      |\varepsilon_i|^2 \indic{\tau_R > n} \bigg]
    \le K \sum_{i = 1}^\infty \frac{\gamma_i^\lambda}{i}
    < \infty
    ~~ \mbox{so} ~~
    \sum_{i = 1}^\infty \frac{1}{i}
    \epsilon_i^2 \indic{\tau_R > n} < \infty
    ~ \mbox{a.s.}
  \end{equation*}
by Condition~\eqref{eqn:lazy-gamma}. 
%  
%  Since
%  \begin{align*}
%\left\{ \sup_i |\Delta_i| \leq R \right\} \cap \left\{ \sum_{i=1}^\infty \frac{|\Delta_i|^{1+\lambda} \indic{\tau_T > n} }{\sqrt{i}} < \infty  \right\}  
%\subset \left\{ \sum_{i=1}^\infty \frac{|\Delta_i|^{1+\lambda}}{\sqrt{i}} < \infty  \right\}
%  \end{align*}
%it follows from \eqref{eq:sup_delta} and by the arbitrary choice of $\epsilon$ that
  %\newtext{As it follows from \cite[Eq.~(A11)]{polyak1992acceleration} that $\Prob\left( \sup |\Delta_t| < \infty  \right)=1$}, we obtain
%As Lemma~\ref{lemma:converging-power-sum} implies in particular that $\{\Delta_i\}$ is a.s. bounded, 
 As in the proof of Theorems~2 and~4 in \cite{polyak1992acceleration}, the Robbins-Siegmund Theorem \cite{robbins1971convergence} applied to the increment of $|\Delta_t|^2$ implies that for every 
  $\epsilon>0$ there exists some $R'>0$ such that
  \begin{align}
  \label{eq:sup_delta}
  \Prob\left(\sup_i |\Delta_i| \le R' \right) \ge 1-\epsilon.
  \end{align} 
Consequently, there exists some $R'' < \infty$ such that $\tau_{R''} = \infty$. 
We obtain that 
 \begin{equation*}
    \sum_{i = 1}^\infty \frac{1}{i}
    \varepsilon_i^2  < \infty
    ~~ \mbox{a.s.}.
  \end{equation*}
  Applying the square integrable martingale convergence
  theorem of \cite[Ex.~5.3.35]{Dembo16}, we have
  \begin{equation*}
    \frac{1}{\sqrt{n}} \sum_{i = 1}^n
    \left(\frac{1}{\psi'(0)} + w_i^n\right) \epsilon_i \cas 0,
  \end{equation*}
  so that both equations~\eqref{eq:PJ_proof1} and~\eqref{eq:PJ_proof2}
  converge almost surely to 0.
\end{proof}
